\textbf{ Matching the forecast horizon with the relevant ecological processes }

### 2019-07-18 doc/manuscript.tex
Rework introduction to center the space-time distinction in ecological forecasting
Since this paper focuses on time-series vs space-for-time approaches this revision
highlights both the distinction between those two approaches and the ideas behind
why the two approaches may behave differently at different time horizons.

A recent flurry of synthesis papers in high impact journals attests to surging interest in ecological forecasting. Many of the papers 
argue that a commitment to ecological forecasting is necessary to guide the environmental policy and management
decisions that will become increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
A second argument is that rigorous testing of predictions will accelerate the process of basic research and discovery 
\citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
Taken together, these papers, along with a new textbook \citep{dietze_ecological_2017} represent an emerging consensus that 
forecasting should be a priority for ecological research.

No such consensus is available for exactly how to tackle this challenge.
Ecologists rely on many different approaches and models for making predictions, and 
frequently argue about the virtues of process-based vs. phenomenological models (refs?).
We often justify investments in basic research by claiming that improved process-level understanding will improve predictions (e.g.,\citealt{Adler2012}),
and it is easy to find examples of how extrapolation of linear relationships leads to poor predictions under novel conditions \citep{Fitzpatrick,Veloz2012}.
But in many cases simple phenomenological models make better predictions than complex process-based models with high parameter
uncertainty (refs? \citealt{tredennick_we_2017}?).

The pragmatic solution to these debates is clear: we should forecast using the models that make the best predictions.
Identifying these models requires rigorously testing predictions against observations. However, many common ecological predictions,
such as a species' geographic range in the year 2100, are virtually impossible to test (but see \citealt{Maguire2015}), 
leading \citep{dietze_iterative_2018} to call for greater emphasis on near-term forecasting. The assumption is that building 
our capacity to make short-term forecasts will ultimately improve our long-term forecasts.

Our goal is to consider a hypothesis that, in our view, has not been carefully considered by either the push for near-term forecasting or
ongoing debates about competing modeling approaches. We expect that short and long-term forecasts will require models
incorporating different kinds of processes and fit to different kinds of data sets. We demonstrate this idea using two simulation case studies, the first 
focused on how competitive interactions affect population dynamics of a focal species, the second for  an 
eco-evolutionary scenario. Our analyses show that: 
\begin{enumerate}
	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow 
	processes such as evolutionary and ecological selection as well as dispersal.
	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while 
	data spanning spatial gradients capture the long-term outcome of selection and dispersal.  Whether predictive models should be trained using longitudinal or
	spatial data sets, or both, depends on the time-scale of the desired forecast.
	\item A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
\end{enumerate}

>>>>

* Forecasting is important
* Two different approaches to this time-series & space-for-time
	* Not clear which is best
	* Some limited comparisons
* Which approach is best may depend on the time-horizon
* Here we explore this possibility using simulation based approach

Forecasting is increasingly recognized as important to the advancement and application of ecological research.
Ecological forecasting is necessary to guide the environmental policy and management
decisions that are increasingly critical under global change \citep{clark_ecological_2001,mouquet_review:_2015,dietze_iterative_2018}.
In addition, evaluating the accuracy of ecological forecasts can provide rigorous tests of predictions from ecological models that can
advance understanding of the processes governing ecological systems  
\citep{houlahan_priority_2017,dietze_prediction_2017,dietze_iterative_2018}.
This combined value of advancing our fundamental understanding of ecology and applying that understanding to support management and decision making 
makes that forecasting is an important priority for ecological research.

Ecological models used for forecasting have traditionally been based on either time-series modeling or space-for-time substitutions.
Time-series approaches use the historical behavior of a system to fit models that describe the temporal dynamics of that system and use those
dynamic models to make predictions about what will happen in the future. This is a common approach for population based forecasts and includes both
process based population models and data-driven models capturing empirical patterns in data (e.g., \citep{ward_complexity_2014}) 
Alternatively, space-for-time modeling approaches develop models of how and ecological property varies in across space in response to environmental
conditions. These spatial relationships are assumed to also hold for changes at a site through time, and forecasts are made by forecasting future
environmental conditions and determining the associated ecological response. Most forecasts for future species distributions and biodiversity
are based on this type of approach, with a models built to describe the environmental conditions where species occur across space being applied to
forecast environmental conditions to determine where they will occur in the future (e.g., citation). Despite both of these approaches being
widely used there has been little comparison of how they compare for forecasting and understanding the dynamics of ecological systems \citep{harris_forecasting_2018}.

Whether historical dynamics, contemporary spatial patterns, or a combination of the two will serve as the best source of information for forecasting
may depend on how far into the future we are attempting to forecast \citep{harris_forecasting_2018}. This potential dependency on the "forecast horizon"; ()\textit{sensu} citation)
results from lags in the response of ecological conditions to environmental change, changes in the importance of ecological processes with
time scale \citep{levin_1992,rosenzweig_1995}, and differences between time-series and spatial gradients in the range of conditions present in the data.
At short forecast horizons (days to years) conditions are likely to stay within the range of the historical data at a site, the time scales of change will be similar to
the time scales present in that historical data, and the current state of the system is likely to capture the influence of unmeasured processes.
As a result, for near-term forecasts time-series approaches may effectively capture the key dynamics and result in accurate forecasts. In contrast,
at long forecast horizons (decades) an ecosystem is likely to experience conditions that have not been historically observed () and be responding to
processes operating at scales difficult to detect in even long ecological time-series. Therefore, for the kinds of far-term forecasts typically conducted
in biodiversity and species-distribution modeling space-for-time based approaches may include important information about how ecosystems respond to major
shifts in climate over long periods and as a results produce better long-term forecasts. Using different approaches for different time horizons is common
in other areas of forecasting, including forecasts for environmental conditions where models for short-term changes (weather) differ substantially from
those for long-term changes (climate).

Here we investigate the possibility that the best approaches for building models for ecological forecasting may depend on the time horizon of the forecast
and explore how time-series and space-for-time approaches might be combined to address mid-term ecological forecasting. We do this using two simulation case studies, one 
focused on how competitive interactions affect population dynamics of a focal species and one eco-evolutionary scenario. Our analyses show that: 
\begin{enumerate}
	\item For short-term forecasts, phenomenological time-series approaches are hard to beat, whereas longer-term forecasts require understanding slow 
	processes such as evolutionary and ecological selection as well as dispersal.
	\item Different kinds of data reflect the operation of different processes: longitudinal data capture autocorrelation and responses of current assemblages, while 
	data spanning spatial gradients capture the long-term outcome of selection and dispersal.  Whether predictive models should be trained using longitudinal or
	spatial data sets, or both, depends on the time-scale of the desired forecast.
	\item A key challenge for future research is determining the rate at which slow processes first influence and ultimately dominate dynamics.
\end{enumerate}
